Intrusion Detection with \\ Genetic Algorithms and Fuzzy Logic

### 2013-11-09 Talk/Presentation.tex
Worked on the fuzzy algorithm slide
Added a different trapezoid figure, also added stuff to the fuzzy genetic algorithm slide.

\subsection{Data Sets - KDD99 and RLD09}
\begin{frame}
  \frametitle{KDD99}
	\begin{itemize}
		\item Generated by simulating a military network environment in 1999.
		\item Has long been a standard data set for intrusion detection.
		\item Data in the set is classified as normal or attack activity.

		\item KDD99 uses 41 features.
		\begin{itemize}
		  	\item \emph{Features} are properties of a \emph{record}, (either an attack or normal activity), that are used to describe the activity.
		\end{itemize}

@@

\begin{frame}
  \frametitle{Some Features of KDD99}
	\begin{enumerate}
		\item duration: length of the normal or attack activity
in seconds.
		\item src\_bytes: number of bytes sent from source to destination.
		\item num\_failed\_logins: number of failed login attempts.
		\item root\_shell: returns 1 if root shell is obtained, else returns 0.
		\item num\_access\_files: number of operations on access control files.
		\item srv\_count: number of connections to the same service as the current connection in the past two seconds.
		\item serror\_rate: percentage of connections that have ``SYN" errors.
		\item same\_srv\_rate: percentage of connections to the same service.
	\end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{RLD09}
	\begin{itemize}
		\item RLD09 was created because KDD99 is 14 years old.
		\item Data was captured from a university in Bangkok, Thailand.
		\item The data has 10 million data packets.
		\item 17 different types of attacks - divided into denial of service attacks, probe attacks. It also has normal activity.
		\item 12 features, which include the number of packets, source ports, and destination ports.
	\end{itemize}
\end{frame}

\subsection{Rules}
\begin{frame}
  \frametitle{Rules}
	\begin{itemize}
		\item Elements of one set are separated into different sets in order to differentiate between normal connections and attacks.

		\item If <\emph{condition}> then <\emph{action}>.

		\item Specify the details of a packet such as the IP address or port number.
		\item If a packet matches any of the rules in the intrusion detection system, the system will take appropriate action, which may include stopping the connection or logging off the system.

\subsection{Fuzzy Logic}
\begin{frame}
  \frametitle{Fuzzy Logic}
  \begin{columns}
  \begin{column}{0.6\textwidth}
  \begin{itemize}
  	\item Used to detect patterns that have a behavior that is between normal and unusual.
  	\item Fuzzy logic rule: If <\emph{condition}> then <\emph{consequence}>.
  	\begin{itemize}
  		\item \emph{condition} is a fuzzy variable
  		\item \emph{consequence} is a fuzzy set
  	\end{itemize}
  \end{itemize}  
  \end{column}

  \begin{column}{0.4\textwidth}
  \includegraphics[width=0.95\textwidth]{../triangleFigure.JPG}
  \end{column}
  \end{columns}
\end{frame}

@@

\subsection{Determining the Accuracy of an Algorithm}
\begin{frame}
  \frametitle{Determining the Accuracy of an Algorithm}
Explain training and test set, false positive, false negative, true positive, true negative, detection rate.
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Genetic Algorithm Implementation]{Genetic Algorithm Implementation}
\subsection{Algorithm Overview}
\begin{frame}
  \frametitle{Algorithm Overview}

@@

\subsection{Main Points of Research}
\begin{frame}
  \frametitle{Main Points of Research}
	\begin{itemize}
		\item Detecting new or unknown types of attacks in a network.
		\item The intrusion detection system used is able to identify normal network activity as well as attacks using a fuzzy genetic algorithm.

		\item Ran experiments using only RLD09, and experiments using KDD99 and RLD09 together.
	\end{itemize}
\end{frame}

@@

  \begin{columns}
  \begin{column}{0.4\textwidth}
  \begin{itemize}
  	\item Trapezoidal shape
  \end{itemize}

  \includegraphics[width=0.95\textwidth]{../trapFigure.jpg}

  \begin{itemize}
	\item The parameters are the values of a feature.
  \end{itemize}

  \end{column}

  \begin{column}{0.6\textwidth}
  \begin{itemize}
  	\item Fuzzy algorithm
  \end{itemize}

\begin{algorithmic}
\IF{data value is between $b$ and $c$}
  \STATE{prob = 1.0}
\ELSIF{data value is between $a$ and $b$}
  \STATE{
    prob = $(\textrm{data}-a)/(b-a)$
  }
\ELSIF{data value is between $c$ and $d$}
  \STATE{
    prob = $(d-\textrm{data})/(d-c)$
  }
\ELSE \STATE{prob = 0.0}
\ENDIF
\end{algorithmic}

@@

\begin{frame}
	\frametitle{Algorithm Overview}

>>>>

\subsection{Data Sets - KDD99 and RLD09}
\begin{frame}
  \frametitle{KDD99}
	\begin{itemize}
		\item Generated by simulating a military network environment in 1999.
		\item Has long been a standard data set for intrusion detection.
		\item Data in the set is classified as normal or attack activity.
		\item KDD99 uses 41 features.
		\begin{itemize}
		  	\item \emph{Features} are properties of a \emph{record}, (either an attack or normal activity), that are used to describe the activity.
		\end{itemize}

@@

\begin{frame}
  \frametitle{Some Features of KDD99}
	\begin{itemize}
		\item duration: length of the normal or attack activity
in seconds.
		%\item src\_bytes: number of bytes sent from source to destination.
		\item num\_failed\_logins: number of failed login attempts.
		\item root\_shell: returns 1 if root shell is obtained, else returns 0.
		%\item num\_access\_files: number of operations on access control files.
		%\item srv\_count: number of connections to the same service as the current connection in the past two seconds.
		\item serror\_rate: percentage of connections that have ``SYN" errors.
		%\item same\_srv\_rate: percentage of connections to the same service.
	\end{itemize}
\end{frame}

\begin{frame}
  \frametitle{RLD09}
	\begin{itemize}
		\item RLD09 was created because KDD99 is 14 years old.
		\item Data was captured from a university in Bangkok, Thailand.
		\item The data has 10 million data packets, 17 different types of attacks (divided into denial of service and probe attacks), and 12 features.
	\end{itemize}
\end{frame}

\subsection{Rules}
\begin{frame}
  \frametitle{Rules}
	\begin{itemize}
		\item Elements of one set are separated into different sets in order to differentiate between normal connections and attacks.
		\item If-Then format
		\begin{itemize}
			\item If the length of the activity is 4 seconds, then the probability of it being an attack is 100\%.
		\end{itemize}				
	\end{itemize}
\end{frame}

@@

\subsection{Determining the Accuracy of an Algorithm}
\begin{frame}
  \frametitle{Determining the Accuracy of an Algorithm}
Explain false positive, false negative, true positive, true negative, detection rate.
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Genetic Algorithm Implementation]{Genetic Algorithm Implementation}
\subsection{Algorithm Overview}
\begin{frame}
  \frametitle{Algorithm Overview}

@@

@@

  \begin{columns}
  \begin{column}{0.6\textwidth}
  \includegraphics[width=0.95\textwidth]{../TrapezoidFigure.pdf}
  \end{column}

  \begin{column}{0.4\textwidth}
  Example:
  \begin{itemize}
  	\item Feature: duration (length of the activity in seconds).
  	\item a=1, b=3, c=5, d=7
  	\item The length of the activity is 6 seconds (between c and d).
  	\linespace
  	\item prob = $\frac{d-\textrm{data}}{d-c}$ = $\frac{7-6}{7-5}$ = 0.5
  \end{itemize}

@@

\begin{frame}
	\frametitle{Algorithm Overview}
\begin{columns}
\begin{column}{0.6\textwidth}
\begin{algorithmic}
\FOR{each record}
  \FOR{each rule}
    \FOR{each feature}
      \STATE{prob = fuzzy(); // Trapezoidal fuzzy rule shape}
      \STATE{totalprob = totalprob + prob;}
    \ENDFOR    
    \IF{totalprob > threshold} 
      \STATE{class is attack;}
      %\ELSE \STATE{class is normal;}
    \ENDIF
  \ENDFOR
  \STATE{find $A$, $B$, $\alpha$, and $\beta$  %// $A$: \# of attack records. $B$: \# of normal records. $\alpha$: \# of attack records correctly identified as attack. $\beta$: \# of normal records incorrectly classified as attack.}

  } 
\ENDFOR
\STATE{calculate fitness}
\STATE{crossover(), mutation()}

\end{algorithmic}
\end{column}

\begin{column}{0.4\textwidth}
	Fitness function:
	\begin{equation*}
	\frac{\alpha}{A} - \frac{\beta}{B}
	\end{equation*}

	$A$: \# of attack records.

	$B$: \# of normal records.

	$\alpha$: \# of attack records correctly identified as attack.

	$\beta$: \# of normal records incorrectly classified as attack.

\begin{frame}
	\frametitle{Experiments}
A variety of experiments were run. Two experiments used just RLD09, and three experiments used KDD99 and RLD09 together.
